---
title: "Links"
---

```{r}
#| echo: false
#| message: false
#| warning: false


library(tidyverse)
```

## Websites

* Tidy Tuesday data repository, link [here](https://github.com/rfordatascience/tidytuesday/tree/master).

* The R Graph Gallery, link [here](https://r-graph-gallery.com/index.html).

* From Data to Viz website on visualizing data, link [here](https://www.data-to-viz.com/#explore).

## Texts

* Spatial Data Science with applications in R available [here](https://r-spatial.org/book/).

* Deep Learning and Scientific Computing with R torch available [here](https://skeydan.github.io/Deep-Learning-and-Scientific-Computing-with-R-torch/).

* Analyzing US Census Data: Methods, Maps, and Models in R textbook available [here](https://walker-data.com/census-r/).

* Tidy Finance with R available [here](https://www.tidy-finance.org/).

* Outstanding User Interfaces with Shiny textbook available [here](https://unleash-shiny.rinterface.com/).

* Crime by the Numbers: A Criminologist’s Guide to R textbook available [here](https://crimebythenumbers.com/).

## Videos

*  Getting started with Quarto YouTube video, link [here](https://youtu.be/YK6IjJkjkiI).
  
```{r}
#| echo: false

vembedr::embed_youtube(id="_f3latmOhew",height=450) %>%
  vembedr::use_align("center")
```

*  Datasheets for Datasets video, link [here](https://youtu.be/R7s7_T4yXak).
  
```{r}
#| echo: false

vembedr::embed_youtube(id="R7s7_T4yXak",height=450) %>%
  vembedr::use_align("center")
```

