---
title: "Lesson 1"
subtitle: "Getting Into Data"
author: "JMG"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    toc: true
    toc-location: left
bibliography: lesson01.bib
---

```{r}
#| echo: true
#| message: false
#| warning: false

# load packages used in document
library(tidyverse)
```


## Learning Objectives

* 


## Readings, etc.

* Read the preface of *Data Science: A First Introduction* by Tiffany Timbers, Trevor Campbell, and Melissa Lee [@timbers2022data]. This textbook is available online [here](https://datasciencebook.ca/).

* Throughout the semester, we will use example data from the [Tidy Tuesday data repository](https://github.com/rfordatascience/tidytuesday) [@TT]. Click [link here](https://github.com/rfordatascience/tidytuesday) to go to the Tidy Tuesday data repository.



## References

::: {#refs}
:::


:::{.callout-tip collapse="true"}
## Expand for Session Info
```{r}
#| echo: false


library(sessioninfo)
# save the session info as an object
pkg_sesh <- session_info(pkgs = "attached")

# get the quarto version
quarto_version <- system("quarto --version", intern = TRUE)

# inject the quarto info
pkg_sesh$platform$quarto <- paste(
  system("quarto --version", intern = TRUE), 
  "@", 
  quarto::quarto_path()
  )

# print it out
pkg_sesh
```

:::




